\noindent{\bf Climate research.}
Climate modeling applications address one of the most
pressing issues faced by our society. Numerous surveys show
increasing trends in the average sea-surface and air temperatures.
A promising method of predicting the future effects of climate
change is to run large-scale simulations of the global climate
many years into the future. In such simulations, climatologists
discretize our entire globe, setting up differential equations to model
Earth's climate. These climate simulations can
generate several terabytes, or even petabytes, of raw data.
Such data-intensive applications present a tremendous set of
challenges for computer centers that strive to accommodate,
maintain, transfer, and preserve data. 

Large-scale computing
centers that host such applications cannot afford to frequently
stall on data accesses and lose precious computing cycles.
The I/O performance of scientific applications such as climate
models has a great impact on the completion of scientific
simulation and post-mortem data visualization and analysis.
However, many applications have complex data characteristics
that are not well supported by existing parallel I/O libraries.
For example, applications may generate large number of small
variables. While the data is large in total, each process only holds very small
amount of data for each variable. It is challenging to provide
good I/O speed for both writing and reading. Moreover,
it is also quite common for data post-processing to examine
data along time dimension, e.g. observing the change of
temperature within one hour. However, storage data layouts currently common in use
do not provide a good support for such access pattern.

NASAs climate and weather models such as GEOS-5 (the
Goddard Earth Observing System Model) and Model-E are
such applications. For example, a GEOS-5 simulation at a coarse resolution of half-degree
generates only 3.12GB data but which consists of 185 2-D variables
and 80 3-D variables at a time. The number of time steps can be configured to the
order of thousands. Writing and reading of such small variables poses great
challenges  on current large-scale storage systems.  

An I/O technique that can
significantly enhance the spatial and temporal resolution of
these climate and weather models can systematically lift the
capability of climatologists and meteorologists in addressing
critical climate issues. An I/O method of ADIOS was specifically designed to address these problems. It leverages the spatial and temporal
relationships between variable data. Spatially it aggregates and merges data chunks of
the same time steps so that fewer processes write
larger blocks.
Moreover, the same variable across different time steps are merged together
with time as a new dimension, which again reduces the amount of I/O
request for both writing and reading. This strategy has provided two orders of magnitude of speedup compared to simple writing (and reading) strategies. 





