%\chapterauthor{Norbert Podhorszki and Scott Klasky and Manish Parashar and Nagiza Samatova and Karsten Schwan and Yuan Tian and Matthew Wolf}{ADIOS}

\chapter{ADIOS}
\label{part3-ch5-adios}
{\color {red}Please target 5-8 pages for this chapter (including figures+references). Readers should be able to appreciate the features of the I/O library, and why they might consider using it for their applications.}


\section{Motivation}
{\color {red}Scott}

\input{Chapters/part3-ch5-adios/intro.tex}


\section{Design/Architecture}
{\color {red}Norbert}

\input{Chapters/part3-ch5-adios/design.tex}


\section{Deployment, Usage, applications}
{\color {red}Brief summary of where the library is deployed, who uses it, etc.

Note: you can choose to mention some performance numbers in the context of applications, but for the purpose of this book we would like you to avoid direct comparisons of your favorite library with other technologies.  }

ADIOS is an open source software with a BSD license...


\subsection{Checkpoint/restart}
{\color {red}Norbert: use RAMGEN as example}
\input{Chapters/part3-ch5-adios/ckp.tex}


\subsection{Analysis}
{\color {red}Karsten/Matt/Norbert: use J.Tromp as example?}
\input{Chapters/part3-ch5-adios/analysis.tex}
{\color {red}Yuan}
\input{Chapters/part3-ch5-adios/analysis-climate.tex}


\subsection{Visualization}
{\color {red}Norbert: use Fusion as example}
\input{Chapters/part3-ch5-adios/visualization.tex}


\subsection{Code coupling}
{\color {red}Manish: use Fusion as example}
\input{Chapters/part3-ch5-adios/coupling.tex}


\subsection{Data reduction}
{\color {red}Nagiza: PARLO}
\input{Chapters/part3-ch5-adios/reduction.tex}



\subsection{Deployment}
ADIOS has been installed as central software and is supported organizationally by the OLCF, by the iVEC organization in Western Australia for the purpose of supporting the Square-Kilometer Array project, by the High Level Support Team (HLST) supporting EFDA (European Fusion Development Agreement) sites, and by the ERDC center of the US. Army Corps of Engineers. At other locations, application scientists install ADIOS themselves for their own application. Many of those receive guidance from the ADIOS team located at the OLCF to utilize their target system optimally.



\index{Data Formats}
\index{I/O middleware}


\section{Conclusion}
{\color {red}Scott}
{\color {red}Summary of experiences with the library. Future challenges and directions.}

%This is a reference to a book chapter~\ref{ch0:book-intro}.
%This is a reference to a bibtex item~\cite{Snir:MPI}.}

\putbib[Chapters/part3-ch5-adios/part3-ch5-adios]
